### 2021-11-11 NSF2021/Description/specificaim2.tex

In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align}
  \label{eq:F}
  F[u] = -\rho^2 \Delta u + f'(u).
\end{align}
%\begin{wrapfigure}[13]{l}{1.8in}
%  \centering
%  \input{figures/doublewell.tikz}
%  \caption{\label{fig:well} $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$.}
%\end{wrapfigure}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - \alpha dF^{-1} F(u^N),
\end{align*}
where $dF = -\rho^2 \Delta + f''(u^N)$ is the Fr\'{e}chet derivative of
$F$, and $\alpha$ is the step size which will be determined using
backtracking.
As expected, the biggest challenge amounts to solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = F[u^N].
\end{align}
\begin{wrapfigure}[20]{r}{1.8in}
  \centering
  \input{figures/doublewell.tikz}  \\
  \input{figures/residuals.tikz}
  \caption{\label{fig:residuals} A double well (top) and the error (bottom)
    for the case $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$.}
\end{wrapfigure}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging. We propose to instead use an inexact
Newton's method where $dF^{-1}$ is defined in terms of the piecewise
quadratic approximation
\begin{align*}
  f(u) \approx \tilde{f}(u) = bu + \begin{cases}
    a(u - U_1)^2, &\mbox{if } u \leq (U_0 + U_1)/2 \\
    (u - U_2)^2, &\mbox{if } u > (U_0 + U_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + \tilde{f}''(u^{N})v = F[u^N],
\end{align}
which is a screened Poisson equation with piecewise a constant screening
coefficient. We test the efficacy of Newton's method applied
to~\eqref{eq:F} by considering its one-dimensional version in $(0,1)$
with the Dirichlet boundary conditions $u(0) = U_0$ and $u(1) = U_1$.
These parameters result in the double well potential illustrated in
Figure~\ref{fig:well}(a), and the inexact Newton's method converges to
the solution in Figure~\ref{fig:well}(b). Figure~\ref{fig:residuals}
shows the max norm of the residual at each Newton iterate when using
the exact second derivative (red curve) and the piecewise constant
approximate (blue curve).

>>>>

In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align}
  \label{eq:F}
  F[u] = -\rho^2 \Delta u + f'(u).
\end{align}
%\begin{wrapfigure}[13]{l}{1.8in}
%  \centering
%  \input{figures/doublewell.tikz}
%  \caption{\label{fig:well} $U_1 = 0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$.}
%\end{wrapfigure}
This non-linear PDE will be solved with Newton's method
\begin{align*}
  u^{N+1} = u^{N} - \alpha (dF^N)^{-1} F(u^N),
\end{align*}
where $dF^N = -\rho^2 \Delta + f''(u^N)$ is the Fr\'{e}chet derivative
of $F$, and $\alpha$ is the step size which will be determined using
backtracking. As expected, the biggest computational challenge is
solving
\begin{align}
  \label{eq:NewtonUpdate}
  - \rho^2 \Delta v + f''(u^{N}) v = F[u^N].
\end{align}
\begin{wrapfigure}[]{l}{1.8in}
  \vspace{-46pt}
  \centering
  \input{figures/doublewell.tikz} 
  \caption{\label{fig:CA} (a) The double well potential $U_1 =
  0.5$, $U_2 = 2$, $a = 1$, $b = 0.4$. (b) The solution of the
  Allen-Cahn equation~\eqref{eq:SL}. (c) The convergence of the exact
  and inexact Newton's methods.}
\end{wrapfigure}
For a general double well potential $f(u)$, this is a variable
coefficient elliptic PDE, and developing an integral equation
formulation is challenging since the fundamental solution is not easily
determined. We propose to instead use an inexact Newton's method where
$dF^{-1}$ is defined in terms of the piecewise quadratic approximation
\begin{align*}
  f(u) \approx \tilde{f}(u) = bu + \begin{cases}
    a(u - U_1)^2, &\mbox{if } u \leq (U_0 + U_1)/2 \\
    (u - U_2)^2, &\mbox{if } u > (U_0 + U_1)/2.
  \end{cases}
\end{align*}
In this case, the update step of Newton's method~\eqref{eq:NewtonUpdate}
amounts to solving the PDE
\begin{align}
  \label{eq:screenedPoisson}
  -\rho^2 \Delta v + \tilde{f}''(u^{N})v = F[u^N],
\end{align}
which is a screened Poisson equation with piecewise a constant screening
coefficient. We test the efficacy of Newton's method applied
to~\eqref{eq:F} by considering a finite difference method applied to its
one-dimensional version in $(0,1)$ with the Dirichlet boundary
conditions $u(0) = U_0$ and $u(1) = U_1$.  These parameters result in
the double well potential illustrated in Figure~\ref{fig:CA}(a), and the
inexact Newton's method converges to the solution in
Figure~\ref{fig:CA}(b). Figure~\ref{fig:CA}(c) shows the max norm of the
residual at each Newton iterate when using the exact second derivative
(red curve) and the piecewise constant approximate (blue curve).
